---
title: "time_to_anoxia_modeling"
author: "Jianqiu Zheng"
date: "3/5/2023"
output: html_document
---

## Building simple kinetic model
## concentration in mg/L
```{r kinetic_model_demo}
library(FME)

solveDOKin <- function(pars, times=seq(0,30, by=0.05)){
    derivs <- function(t, state, pars){
      with(as.list(c(state, pars)),{
      r1<- vmax1* Oxy/(Oxy+10)*CHO/(CHO+10)  #dual-substrate   
      dOxy <--r1*BM1  #stoichiometry
      dCHO <--r1*BM1
      dBM1 <- r1*BM1
    return(list(c(dOxy, dCHO,dBM1)))
      })
  }
times=seq(0,30, by=0.05)
state<-c(Oxy=8,CHO=50, BM1=1)
#solved the model
  ox<-ode(y=state, times=times, func=derivs, parms=pars)

data.frame(ox[,1:4])
}

pars1<- list (vmax1=1)
output1<-solveDOKin(pars1)  


##plotting
#pdf("DO_dynamic.pdf", width=3.4, height=3.6)
plot(output1$time,output1$Oxy, type = "l", main="Microbial uptake rate",xlab = "Time (h)", ylab = "DO (mg/L)", lwd=3,lty=1,col="#fdb863", xlim=c(0,30), ylim=c(0,10))

#dev.off()


```


# Fitting observational data with linear model (abiotic)
```{r abiotic}

raw<-read.csv("DO_dynamic_modeling/wetlandA.csv",header=TRUE)
sub1<-raw[raw$sample_name =='anoxia_062',]
Data<-sub1[, c("time_hour","DO_corrected_mgL")]
colnames(Data)<-c("time","Oxy")


solveABO <- function(pars, times=seq(0,30, by=0.05)){
    derivs <- function(t, state, pars){
      with(as.list(c(state, pars)),{
      dOxy <--va*Oxy  #
    return(list(c(dOxy)))
      })
  }
times=seq(0,30, by=0.05)
state<-c(Oxy=5.2) #mg/g
#solved the mode
  ox<-ode(y=state, times=times, func=derivs, parms=pars)

data.frame(ox[,1:2])
}

pars<- list (va=1)
#output1<-solveLIN(pars1)  


Objective <- function(x, parset = names(x)) {
 pars[parset] <- x
 tout <- seq(0, 30, by = 0.05) ## output times
 out <- solveABO(pars, tout)
 return(modCost(model = out, obs = Data))## Model cost
}

#parameters are constrained to be >0
print(system.time(Fit <- modFit(p = c(va=1),
 f = Objective, lower = c(0.0))))

summary(Fit)
result<-list(Fit$par, Fit$ssr)
result

##plotting of single fitting
pars[c("va")]<-Fit$par
#pars<- list (vmax1=1,vmax2=46,vmax3=69)
output<-solveABO(pars)  

mfit<-as.data.frame(output[, 1:2])
colnames(mfit) <- c("time", "Oxy")
head(mfit)


#pdf("lowC_highN.pdf", width=6, height=4)
#par(xpd = T, mar = par()$mar + c(0,0,0,10))
plot(Data$time,Data$Oxy, xlab = "Time (d)", ylab = "Concentration (uM)", pch=16, cex=1.2,col="#e08214", xlim=c(0,30), ylim=c(0,10),cex.lab=1.2, cex.axis=1.2)
lines(output$time, output$Oxy, lwd=4, col="black")
#legend("topright",inset = c(- 0.6, 0), c("NO3","NO2","Inflow"), lwd=2, lty=c(1,3),col=c("#e08214","#e08214","#e08214"))

#dev.off()

#--------plotting all reps
pars1<- list (va=0.209)
output1<-solveABO(pars1)

pars2<- list (va=0.4579)
output2<-solveABO(pars2) 

pars3<- list (va=0.5668)
output3<-solveABO(pars3)  


final<-cbind(output1$time, output1$Oxy,output2$Oxy, output3$Oxy)
colnames(final)<-c("time","DO1","DO2","DO3") 
write.csv(final, "wetA_ABO.csv")


raw<-read.csv("DO_dynamic_modeling/wetlandA.csv",header=TRUE)
pdf("wetA_ABO.pdf", width=3.4, height=3.6)
plot(raw$time_hour,raw$DO_corrected_mgL, xlab = "Time (h)", ylab = "DO (mg/L)", pch=1, cex=1.2,col="#fdb863", xlim=c(0,25), ylim=c(0,10),cex.lab=1.2, cex.axis=1.2)
lines(output1$time, output1$Oxy, lty=3,lwd=2, col="#7f3b08")
lines(output2$time, output2$Oxy, lty=3,lwd=2, col="#7f3b08")
lines(output3$time, output3$Oxy, lty=3,lwd=2, col="#7f3b08")
legend("topright", c("Obv","Model"), pch=c(1,NA), lty=c(0,3),lwd=c(NA, 2), col=c("#fdb863","#7f3b08","#b35806","#e08214","#fdb863","#542788","#8073ac"))

dev.off()


```



## fitting observational data with nonlinear model. Note: 3 replicates for each soil, replicates are fitted independently
```{r data_fitting}
##load data
raw<-read.csv("DO_dynamic_modeling/uplandA.csv",header=TRUE)
sub1<-raw[raw$sample_name =='anoxia_047',]
Data<-sub1[, c("time_hour","DO_corrected_mgL")]
colnames(Data)<-c("time","Oxy")

Data$diff<-(Data$Oxy-mean(Data$Oxy))^2
TSS<-sum(Data$diff)

#function
library(FME)

solveMM <- function(pars, times=seq(0,30, by=0.05)){
    derivs <- function(t, state, pars){
      with(as.list(c(state, pars)),{
      r1<- vmax1* Oxy/(Oxy+10) #     
      dOxy <--r1*BM1 #
      dCHO <--r1*BM1
      dBM1 <- r1*BM1
    return(list(c(dOxy, dCHO,dBM1)))
      })
  }
times=seq(0,30, by=0.05)
state<-c(Oxy=8.39,CHO=51.1, BM1=1) #mg/g
#solved the mode
  ox<-ode(y=state, times=times, func=derivs, parms=pars)

data.frame(ox[,1:4])
}


#fitting
Objective <- function(x, parset = names(x)) {
 pars[parset] <- x
 tout <- seq(0, 30, by = 0.05) ## output times
 out <- solveMM(pars, tout)
 return(modCost(model = out, obs = Data))## Model cost
}


#parameters are constrained to be >0
#print(system.time(Fit <- modFit(p = c(vmax1=0.5),
# f = Objective, lower = c(0.0))))

summary(Fit)
result<-list(Fit$par, Fit$ssr)
result

##plotting
pars[c("vmax1")]<-Fit$par
output<-solveMM(pars)  

mfit<-as.data.frame(output[, 1:2])
colnames(mfit) <- c("time", "Oxy")
head(mfit)


#pdf("lowC_highN.pdf", width=6, height=4)
#par(xpd = T, mar = par()$mar + c(0,0,0,10))
plot(Data$time,Data$Oxy, xlab = "Time (d)", ylab = "Concentration (uM)", pch=16, cex=1.2,col="#e08214", xlim=c(0,30), ylim=c(0,10),cex.lab=1.2, cex.axis=1.2)
lines(output$time, output$Oxy, lwd=4, col="black")
#legend("topright",inset = c(- 0.6, 0), c("NO3","NO2","Inflow"), lwd=2, lty=c(1,3),col=c("#e08214","#e08214","#e08214"))

#dev.off()

##plotting all replicates
pars1<- list (vmax1=0.8005)
output1<-solveMM(pars1)

pars2<- list (vmax1=0.6398)
output2<-solveMM(pars2) 

pars3<- list (vmax1=0.4713)
output3<-solveMM(pars3)  

final<-cbind(output1$time, output1$Oxy,output2$Oxy, output3$Oxy)
colnames(final)<-c("time","DO1","DO2","DO3") 
write.csv(final, "upA_MM.csv")

raw<-read.csv("DO_dynamic_modeling/uplandA.csv",header=TRUE)
#pdf("upA_MM.pdf", width=3.4, height=3.6)
plot(raw$time_hour,raw$DO_corrected_mgL, xlab = "Time (h)", ylab = "DO (mg/L)", pch=1, cex=1,col="#fdb863", xlim=c(0,25), ylim=c(0,10),cex.lab=1.2, cex.axis=1.2)
lines(output1$time, output1$Oxy, lty=2, lwd=2, col="#7f3b08")
lines(output2$time, output2$Oxy, lty=2, lwd=2, col="#7f3b08")
lines(output3$time, output3$Oxy, lty=2, lwd=2, col="#7f3b08")
#legend("topright", c("Obv","Model"), pch=c(1,NA), lty=c(0,2),lwd=c(NA, 2), col=c("#fdb863","#7f3b08","#b35806","#e08214","#fdb863","#542788","#8073ac"))

#dev.off()


```


# fitting combined linear + MM formulation
```{r abiotic2}

raw<-read.csv("DO_dynamic_modeling/wetlandA.csv",header=TRUE)
sub1<-raw[raw$sample_name =='anoxia_060',]
Data<-sub1[, c("time_hour","DO_corrected_mgL")]
colnames(Data)<-c("time","Oxy")



solveCOM <- function(pars, times=seq(0,30, by=0.05)){
    derivs <- function(t, state, pars){
      with(as.list(c(state, pars)),{
      r1<- vmax1* Oxy/(Oxy+10) #     
      dOxy <--r1*BM1-va*Oxy  #
      dCHO <--r1*BM1
      dBM1 <- r1*BM1
    return(list(c(dOxy, dCHO,dBM1)))
      })
  }
times=seq(0,30, by=0.05)
state<-c(Oxy=5.97,CHO=51.1, BM1=1) #mg/g
#solved the mode
  ox<-ode(y=state, times=times, func=derivs, parms=pars)

data.frame(ox[,1:4])
}

pars<- list (vmax1=10, va=1)
#output1<-solveLIN(pars1)  


Objective <- function(x, parset = names(x)) {
 pars[parset] <- x
 tout <- seq(0, 30, by = 0.05) ## output times
 out <- solveCOM(pars, tout)
 return(modCost(model = out, obs = Data))## Model cost
}


#parameters are constrained to be >0
print(system.time(Fit <- modFit(p = c(vmax1=0.1, va=0.1),
 f = Objective, lower = c(0.01, 0.0),upper = c(2, 0.5))))

summary(Fit)
result<-list(Fit$par, Fit$ssr)
result

##plotting 
pars[c("vmax1", "va")]<-Fit$par
#pars<- list (vmax1=1,vmax2=46,vmax3=69)
output<-solveCOM(pars)  

mfit<-as.data.frame(output[, 1:4])
colnames(mfit) <- c("time", "Oxy","CHO","Amm")
head(mfit)

plot(Data$time,Data$Oxy, xlab = "Time (d)", ylab = "Concentration (uM)", pch=16, cex=1.2,col="#e08214", xlim=c(0,30), ylim=c(0,10),cex.lab=1.2, cex.axis=1.2)
lines(output$time, output$Oxy, lwd=4, col="black")

#------------------plotting all reps
pars1<- list (vmax1=0.447, va=0.154)
output1<-solveCOM(pars1)

pars2<- list (vmax1=0.101, va=0.449)
output2<-solveCOM(pars2) 

pars3<- list (vmax1=0.0126,va=0.4569)
output3<-solveCOM(pars3)  


final<-cbind(output1$time, output1$Oxy,output2$Oxy, output3$Oxy)
colnames(final)<-c("time","DO1","DO2","DO3") 
write.csv(final, "wetA_comb.csv")


raw<-read.csv("DO_dynamic_modeling/wetlandA.csv",header=TRUE)
pdf("wetA_COMB.pdf", width=3.4, height=3.6)

plot(raw$time_hour,raw$DO_corrected_mgL, xlab = "Time (h)", ylab = "DO (mg/L)", pch=1, cex=1.2,col="#fdb863", xlim=c(0,25), ylim=c(0,10),cex.lab=1.2, cex.axis=1.2)
lines(output1$time, output1$Oxy, lwd=2, col="#7f3b08")
lines(output2$time, output2$Oxy, lwd=2, col="#7f3b08")
lines(output3$time, output3$Oxy, lwd=2, col="#7f3b08")
legend("topright", c("Obv","Model"), pch=c(1,NA), lty=c(0,1), col=c("#fdb863","#7f3b08","#b35806","#e08214","#fdb863","#542788","#8073ac"))

dev.off()


```







### Varying rate ratio
```{r data}

solveCOM <- function(t, state, pars){
      with(as.list(c(state, pars)),{
      r1<- 0.5* Oxy/(Oxy+10)*CHO/(CHO+10)  #denitrification     
      dOxy <--r1*BM1-alpha*0.5*Oxy  #stoichiometry from van den Berg
      dCHO <--r1*BM1
      dBM1 <- r1*BM1
    return(list(c(dOxy, dCHO,dBM1)))
      })
    }


times=seq(0,30, by=0.05)
state<-c(Oxy=8,CHO=50, BM1=1)

pars1=list (alpha=0)
pars2=list (alpha=0.1)
pars3=list (alpha=0.2)
pars4=list (alpha=0.5)
pars5=list (alpha=1)
pars6=list (alpha=2)
pars7=list (alpha=5)

ox1<-ode(y=state, times=times, func=solveCOM, parms=pars1)
output1<-data.frame(ox1[,1:4])


ox2<-ode(y=state, times=times, func=solveCOM, parms=pars2)
output2<-data.frame(ox2[,1:4])


ox3<-ode(y=state, times=times, func=solveCOM, parms=pars3)
output3<-data.frame(ox3[,1:4])


ox4<-ode(y=state, times=times, func=solveCOM, parms=pars4)
output4<-data.frame(ox4[,1:4])


ox5<-ode(y=state, times=times, func=solveCOM, parms=pars5)
output5<-data.frame(ox5[,1:4])

ox6<-ode(y=state, times=times, func=solveCOM, parms=pars6)
output6<-data.frame(ox6[,1:4])


ox7<-ode(y=state, times=times, func=solveCOM, parms=pars7)
output7<-data.frame(ox7[,1:4])



##plotting
pdf("Varying_ratio.pdf", width=3.4, height=3.6)
plot(output1$time,output1$Oxy, type = "l", main="rate ratio",xlab = "Time (h)", ylab = "DO (mg/L)", lwd=3,lty=1,col="#7f3b08", xlim=c(0,30), ylim=c(0,10))
lines(output2$time,output2$Oxy, lwd=3, lty=1, col="#b35806")
lines(output3$time,output3$Oxy, lwd=3, lty=1, col="#e08214")
lines(output4$time,output4$Oxy, lwd=3, lty=1, col="#fdb863")
lines(output5$time,output5$Oxy, lwd=3, lty=1, col="#b2abd2")
lines(output6$time,output6$Oxy, lwd=3, lty=1, col="#8073ac")
lines(output7$time,output7$Oxy, lwd=3, lty=1, col="#542788")

#legend("topright", c("1","2","3","4","5","6","7"), lwd=3, lty=1, col=c("#7f3b08","#b35806","#e08214",  "#fdb863","#b2abd2","#8073ac","#542788"))

dev.off()


```


### Varying rate ratio
```{r data-}

##combined formulation

solveCOM <- function(t, state, pars){
      with(as.list(c(state, pars)),{
      r1<- vmax1* Oxy/(Oxy+10)*CHO/(CHO+10)  #denitrification     
      dOxy <--r1*BM1-0.1*Oxy  #stoichiometry from van den Berg
      dCHO <--r1*BM1
      dBM1 <- r1*BM1
    return(list(c(dOxy, dCHO,dBM1)))
      })
    }

pars=list (vmax1=0.5)
times=seq(0,30, by=0.05)

state1<-c(Oxy=8,CHO=25, BM1=1)
ox1<-ode(y=state1, times=times, func=solveCOM, parms=pars)
output1<-data.frame(ox1[,1:4])

state2<-c(Oxy=8,CHO=37.5, BM1=1)
ox2<-ode(y=state2, times=times, func=solveCOM, parms=pars)
output2<-data.frame(ox2[,1:4])

state3<-c(Oxy=8,CHO=50, BM1=1)
ox3<-ode(y=state3, times=times, func=solveCOM, parms=pars)
output3<-data.frame(ox3[,1:4])

state4<-c(Oxy=8,CHO=62.5, BM1=1)
ox4<-ode(y=state4, times=times, func=solveCOM, parms=pars)
output4<-data.frame(ox4[,1:4])

state5<-c(Oxy=8,CHO=75, BM1=1)
ox5<-ode(y=state5, times=times, func=solveCOM, parms=pars)
output5<-data.frame(ox5[,1:4])



##plotting
pdf("Varying_DOC_COM.pdf", width=3.4, height=3.6)
plot(output1$time,output1$Oxy, type = "l", main="Initial DOC",xlab = "Time (h)", ylab = "DO (mg/L)", lwd=3,lty=1,col="#fdb863", xlim=c(0,25), ylim=c(0,10))
lines(output2$time,output2$Oxy, lwd=3, lty=1, col="#e08214")
lines(output3$time,output3$Oxy, lwd=3, lty=2, col="#5e5e60")
lines(output4$time,output4$Oxy, lwd=3, lty=1, col="#b35806")
lines(output5$time,output5$Oxy, lwd=3, lty=1, col="#7f3b08")
legend("topright", c("-50%","-25%","Ref","+25%","+50%"),lwd=3, lty=c(1,1,2,1,1), col=c("#fdb863","#e08214","#5e5e60","#b35806","#7f3b08"))

dev.off()


```


```{r data--}

solveNR <- function(pars, times=seq(0,400, by=0.1)){
    derivs <- function(t, state, pars){
      with(as.list(c(state, pars)),{
      r1<- vmax1* Oxy/(Oxy+100) * CHO/(CHO+1)  #stoichi: 0.31~0.8    
      dOxy <--r1*BM1  #stoichiometry from van den Berg
      dCHO <--2.7*r1*BM1
      dBM1 <- r1*BM1
    return(list(c(dOxy, dCHO,dBM1)))
      })
  }
times=seq(0,400, by=0.1)
state<-c(Oxy=8,CHO=20, BM1=1)
#solved the mode
  ox<-ode(y=state, times=times, func=derivs, parms=pars)

data.frame(ox[,1:4])
}

pars1<- list (vmax1=1.5)
output1<-solveNR(pars1) 


pars2<- list (vmax1=1.5)
output2<-solveNR(pars2)  

pars3<- list (vmax1=1.5)
output3<-solveNR(pars3) 


#output1,2,3====CHO 10,20,30
pdf("varyingC.pdf", width=4, height=4)
plot(output1$time/24,output1$Oxy, lwd=3, lty=1,type = "l", xlab = "Time (d)", ylab = "DO (mg/L)", col="#fdb863", xlim=c(0,14), ylim=c(0,10))
lines(output2$time/24,output2$Oxy, lwd=3, lty=1, col="#e08214")
lines(output3$time/24,output3$Oxy, lwd=3, lty=1, col="#b35806")

legend("topright", c("DOC=10 mg/L","DOC=15 mg/L","DOC=20 mg/L"), lwd=c(3,3,3),lty=c(1,1,1), col=c("#fdb863","#e08214","#b35806"))
dev.off()


```




## Varying initial DOC concentration (to mimic horizon B)
```{r data---}


derivs <- function(t, state, pars){
      with(as.list(c(state, pars)),{
      r1<- vmax1* Oxy/(Oxy+100) * CHO/(CHO+1)  #stoichi: 0.31~0.8    
      dOxy <--r1*BM1  #stoichiometry from van den Berg
      dCHO <--2.67*r1*BM1
      dBM1 <- r1*BM1
    return(list(c(dOxy, dCHO,dBM1)))
      })
    }

times=seq(0,400, by=0.1)
pars<- list (vmax1=1.5)
state1<-c(Oxy=8,CHO=10, BM1=10)
state2<-c(Oxy=8,CHO=15, BM1=10)
state3<-c(Oxy=8,CHO=20, BM1=10)

ox1<-ode(y=state1, times=times, func=derivs, parms=pars)
output1<-data.frame(ox1[ ,1:4])

ox2<-ode(y=state2, times=times, func=derivs, parms=pars)
output2<-data.frame(ox2[ ,1:4])

ox3<-ode(y=state3, times=times, func=derivs, parms=pars)
output3<-data.frame(ox3[ ,1:4])



#output1,2,3====CHO 10,20,30
#pdf("varyingC.pdf", width=4, height=4)
plot(output1$time/24,output1$Oxy, lwd=3, lty=1,type = "l", xlab = "Time (d)", ylab = "DO (mg/L)", col="#fdb863", xlim=c(0,14), ylim=c(0,10))
lines(output2$time/24,output2$Oxy, lwd=3, lty=1, col="#e08214")
lines(output3$time/24,output3$Oxy, lwd=3, lty=1, col="#b35806")

legend("topright", c("10 mg/L","15 mg/L","20 mg/L"), lwd=c(3,3,3),lty=c(1,1,1), col=c("#fdb863","#e08214","#b35806"))
#dev.off()


```



## box plot--summary of different model fittings
```{r fitted}
library(ggplot2)
library(reshape2)
library(ggpubr)
#theme_pubr()

###biotic rate
up<-c(0.8005,0.6398,0.4713)
#tra<-c(0.7718,0.9996,0.8584)
wet<-c(0.9833,2.4322,2.927)


df<-data.frame(up, wet)
colnames(df)<-c("Upland","Wetland")



data<-reshape2::melt(df)
scaleFUN <- function(x) sprintf("%.2f", x)
p1<-ggplot(data, aes(x=variable, y=value, color=variable))+geom_boxplot()+ scale_color_manual(values=c("#e08214", "#b35806","#7f3b08"))+scale_fill_manual(values=c("#e08214", "#b35806","#7f3b08"))+ylim(0,3)+ylab("Fitted rate (mg/L/h)")+scale_y_continuous(labels=scaleFUN)+theme_pubr(border=TRUE)+theme(panel.grid.minor.x=element_line(linetype="dashed", color="gray",size=0.2))+theme(panel.grid.major.x=element_line(linetype="dashed", color="gray",size=0.2))+theme(axis.title.x=element_blank())+theme(legend.position="none")
p1

pdf("bio_rate_2loc.pdf", width=2.6, height=1.8)
p1
dev.off()


#-----abiotic rate-----
up<-c(0.1884,0.1452,0.1039)
tra<-c(0.1768,0.2155,0.2010)
wet<-c(0.209,0.4579,0.5668)

df<-data.frame(up,  wet)
colnames(df)<-c("Upland","Wetland")



data<-reshape2::melt(df)

p1<-ggplot(data, aes(x=variable, y=value, color=variable))+geom_boxplot(aes(fill=variable),alpha=0.4)+ scale_color_manual(values=c("#e08214", "#b35806","#7f3b08"))+scale_fill_manual(values=c("#e08214", "#b35806","#7f3b08"))+ylim(0,1)+ylab("Fitted rate (mg/L/h)")+theme_pubr(border=TRUE)+theme(panel.grid.minor.x=element_line(linetype="dashed", color="gray",size=0.2))+theme(panel.grid.major.x=element_line(linetype="dashed", color="gray",size=0.2))+theme(axis.title.x=element_blank())+theme(legend.position="none")
p1

pdf("abio_rate2loc.pdf", width=2.6, height=1.8)
p1
dev.off()



#-----------Combined
#abiotic rate
up<-c(0.5455,0.4048,0.3678)
tra<-c(0.588,0.7321,0.6373)
wet<-c(0.4467,0.101,0.0126)

df<-data.frame(up,  wet)
df$rate<-c('vb')
colnames(df)<-c("Upland","Wetland","rate")

#biotic rate
up<-c(0.09667,0.0845,0.0418)
tra<-c(0.0743,0.095,0.0912)
wet<-c(0.1543,0.4496,0.4569)

df2<-data.frame(up,  wet)
df2$rate<-c('vx')
colnames(df2)<-c("Upland","Wetland","rate")

data<-rbind(df, df2)



df<-reshape2::melt(data, id.var='rate')

p1<-ggplot(df, aes(x=variable, y=value, color=variable))+geom_boxplot(aes(fill=rate, alpha=rate))+ scale_color_manual(values=c("#e08214", "#b35806","#7f3b08"))+scale_fill_manual(values=c("#e08214", "#b35806","#7f3b08"))+scale_alpha_discrete(range=c(0,0.4))+ylim(0,1)+ylab("Fitted rate (mg/L/h)")+theme_pubr(border=TRUE)+theme(panel.grid.minor.x=element_line(linetype="dashed", color="gray",size=0.2))+theme(panel.grid.major.x=element_line(linetype="dashed", color="gray",size=0.2))+theme(axis.title.x=element_blank())+theme(legend.position="none")
p1

pdf("comb_rate_2loc.pdf", width=2.6, height=1.8)
p1
dev.off()

```
